---
title: "HW 04"
format: html
editor: visual
embed-resources: true
---

## Set Up

```{r}
#| message: false
library(tidyverse)
```

## Question 1

```{r}
nfl_salaries <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/nfl_salaries.csv")

glimpse(nfl_salaries)
```

## Question 2

Reshape it so that every row tells you the salary associated with a position each year. Your new dataset should have three columns: `year`, `position`, `salary`.

```{r}
tidy_nfl_salaries <- nfl_salaries |>
  pivot_longer(cols = c(2:11),
               names_to = "Position",
               values_to = "Salary")
tidy_nfl_salaries
```

## Question 3

Filter your newly created dataset so that it only contains quarterbacks. Then, make a histogram where salary is in the x-axis. Then use `facet_wrap` to get the histogram for each `year`.

```{r}
tidy_quarterbacks <- tidy_nfl_salaries %>% filter(Position == "Quarterback")
```

```{r}
tidy_quarterbacks <- tidy_quarterbacks |> 
  mutate(new_salary = Salary / 1000000)
```

```{r}
ggplot(tidy_quarterbacks,
       aes(x = new_salary)) +
  geom_histogram()+
  theme_minimal()+
  labs(x = "Salary (in millions)", y = "Count", title = "Quarterback Salary")

```

Then use `facet_wrap` to get the histogram for each `year`.

```{r}
 ggplot(tidy_quarterbacks,
       aes(x = new_salary)) +
  geom_histogram(colour = "black",
                 fill="white")+
  facet_wrap(~year)+
  theme_minimal()+
  labs(x = "Salary (in millions)", y = "Count", title = "Quarterback Salary by Year")
```

**What patterns do you notice?**

There is a higher count of Quarterbacks making less than 1 million. As the years go on, there are more outliers of quarterbacks with higher and higher salaries.

## Question 4

Let's calculate the average salary for each position, each year. Create a new dataset that contains the average salary for each position each year. To do this, you will need the `group_by` and `summarize` combo.

```{r}
salary_position <- tidy_nfl_salaries |> 
  group_by(Position, year) |> 
  summarize(avg_salary = mean(Salary, na.rm = TRUE)) |>
  ungroup()


salary_position
```

## Question 5

Make a linegraph that traces the evolution of each position's average salary across the years. You can use use different strategies to distinguish between positions - color or facets for example. What is important is that you see each position's trajectory clearly and that they are comparable.

```{r}
salary_position <- salary_position |>
    mutate(avg_salary = avg_salary / 1000000)
```

```{r}
ggplot(salary_position,
       aes(x = year,
           y = avg_salary,
           label = Position,
           colour = Position)) +
  geom_line()+
  geom_point(alpha=.5)+
  labs(y = "Salary (in millions)", x = "Year", title = "Change in Salary by Position")
  

```

**Describe at least two trends that are apparent to you.**

The offensive and defensive lineman seem to have steeper salary increases in 2016 through 2018. Throughout all the years, the special team position has the lowest salary.
