---
title: "HW 06"
author: "Kristy"
format: html
editor: visual
embed-resources: true
---

## Question 1

```{r}
library(tidyverse)

# Read in the data 
tv_ratings <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/tv_ratings.csv")

# Glimpse the data 
glimpse(tv_ratings)
```

The following bit of code counts how many seasons each show has and then keeps those shows that have had 5 seasons or more.

```{r}
tv_long <- tv_ratings %>% 
  group_by(title) %>% 
  summarise(num_seasons = n()) %>% 
  ungroup() %>% 
  left_join(tv_ratings, by = "title") 

tv_long <- tv_long %>% 
  filter(num_seasons >= 5)
```

Use `tv_long` to make a line plot to trace how average ratings evolve across seasons. Remember to use the group aesthetic so that each line represents one show.

```{r}
ggplot(tv_long,
             aes(x = seasonNumber,
                 y = av_rating,
                group = title,
                colour = title)) +
  geom_line()+
  labs(title = "Average Ratings Across Seasons",
       x = "Season",
       y = "Average Rating",
       color = "TV Show") +
  theme_minimal()+
  theme(legend.position = "none")
  
```

**It should look fairly messy. Can you draw any conclusions from it?**

Not really.

## Question 2

Facet the plot above by `genres` so that we can see patterns more clearly.

```{r}
ggplot(tv_long,
             aes(x = seasonNumber,
                 y = av_rating,
                group = title,
                colour = title)) +
  geom_line()+
  labs(title = "Average Ratings Across Seasons",
       x = "Season",
       y = "Average Rating",
       color = "TV Show") +
  facet_wrap(~genres) +
  theme_minimal()+
  theme(legend.position = "none")

```

**What shows tend to last longer? Do ratings change much across seasons? Can you identify that show on `Drama, Family, Fantasy` whose ratings just plummeted?**

"Crime, Drama, Mystery" lasted the longest. Ratings do not change much. The show whose ratings just plummeted is "Are you afraid of the dark"

## Question 3

Let's look at the `genres` that show up in the best rated shows.

First, filter the original data set - `tv_ratings` - so that it only includes rows where the average rating is higher or equal than 9.

```{r}
edit_tv_ratings <- tv_ratings %>% 
  filter(av_rating >= 9)
```

Make a barplot where the x-axis is `genre`.

```{r}
ggplot(edit_tv_ratings, 
       aes(x=genres))+
  geom_bar()+
  coord_flip()
```

**The result should be hard to read - the names of the genres are quite long. Add `coord_flip()` to the end of your ggplot call and watch the results get clearer. Tell me what `coord_flip()` does.**

`coord_flip()` switches the x and y axis

**What is the genre with the most top-rated shows?**

Drama

## Question 4

First, because the `genres` names are so convoluted, I am going to classify everything that includes the word "Comedy" as a comedy.

```{r}
comedies_dramas <- tv_ratings %>% 
  mutate(is_comedy = if_else(str_detect(genres, "Comedy"), 
                             1, 
                             0)) %>% # If it contains the word comedy then 1, else 0
  filter(is_comedy == 1 | genres == "Drama") %>% # Keep comedies and dramas
  mutate(genres = if_else(genres == "Drama", # Make it so that we only have those two genres
                          "Drama", 
                          "Comedy"))

glimpse(comedies_dramas)
```

Plot a density plot that shows the distribution of average ratings for both comedies and dramas.

```{r}
s1 <- ggplot(comedies_dramas,
             aes(x = av_rating,
                 y = genres))+
  geom_jitter(alpha = .2)

s2 <-  ggplot(comedies_dramas,
             aes(x = av_rating,
                 y = genres))+
  geom_bin_2d(binwidth = c(.5, .5))

s3 <- ggplot(comedies_dramas,
             aes(x = av_rating,
                 group = genres,
                 colour = genres))+
  geom_density(alpha = .2)

s1 + s2 + s3
```

**How does my prediction above hold? Are dramas rated higher?**

Prof said to keep the first two plots. Comedy has a higher density. But dramas has a higher average rating.

## **Question 5**

First, do the same plot using histograms.

What additional information does this give you?

```{r}
ggplot(comedies_dramas,
             aes(x = av_rating,
                 group = genres,
                 colour = genres))+
  geom_histogram(alpha = .2)
```

Now, use `geom_freqpoly()`. What do you notice? Of the three plots, which one do you think it's more informative?

```{r}
ggplot(comedies_dramas,
             aes(x = av_rating,
                 group = genres,
                 colour = genres))+
  geom_freqpoly()
```

I think the `Freqpoly()` graph is the most informative. It doesn't group average ratings into bars like a histogram would.

## **Question 6**

Let's now explore whether the actual quality of the show corresponded to how many people were paying attention. The column `share` indicates the share of viewership that that show and season acquired. We are going to examine the relationship between average rating and share of viewership.

Take our `comedies_dramas` dataset and make a plot where average rating is on the x-axis and share on the y-axis. Use `geom_bin_2d()` to make the plot.

```{r}
ggplot(comedies_dramas,
             aes(x = av_rating,
                 y = share))+
  geom_bin_2d(binwidth = c(.5, .5))
```

**What do you see? What additional information does this give you in comparison to a scatter plot?**

The average rating is concentrated around 8. A scatter plot shows were the concentration is most dense but a density plot associates a count with the density. We know there are more than 100 count around 8 and below share a of 1.

**Now add `genres` to the fill aesthetic. What patterns do you see? Can you identify that big outlier that apparently captured the nation?**

```{r}
ggplot(comedies_dramas,
             aes(x = av_rating,
                 y = share,
                 fill = genres))+
  geom_bin_2d(binwidth = c(.5, .5))
```

On average comedies have a higher share than dramas. The big outlier was a drama named "Dekalog".
